\documentclass[12pt,a4paper,headsepline,footsepline,DIV13,BCOR12mm]{scrbook}
\usepackage[english,ngerman]{babel}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{scrpage2}
\pagestyle{scrheadings}
\usepackage{setspace}
\usepackage[font=singlespacing]{caption}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{listings}
\usepackage{comment}

\lstset{ % default settings for listings from http://en.wikibooks.org/wiki/LaTeX/Packages/Listings
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true                % sets automatic line breaking
}

%create pdf hyperlinks with a color box (will not be printed)
\usepackage{hyperref}
\hypersetup{colorlinks=false}

% todo and review commands
\usepackage{soul}
\usepackage{color}
%\newcommand{\hl}[1]{\fcolorbox{red}{white}{#1}} %fcolorbox does not break lines properly
\newcommand{\todo}[1]{{\color{red}{\hl{TODO: #1...}}}}
\newcommand{\review}[0]{\todo{REVIEW}}
\newcommand{\comm}[1]{\textsuperscript{\bf \color{red}{\tiny [#1]}}}
\newcommand{\q}[0]{\comm{?}}
\newcommand{\s}[0]{\comm{*}}


\begin{document}

% begin preface
\selectlanguage{ngerman}
\input{tex/titlepage} 
\newpage

\renewcommand{\baselinestretch}{2} % zeilenabstand

%select english as language!
\selectlanguage{english}

\pagenumbering{roman}
\chapter*{Acknowledgements}
I would like to thank everyone who supported me in the last months either by providing professional advice and guidance or just being a good friend and listen to my ideas and thoughts. This work would have never been possible without the support from Prof. Dr. Wolfgang Reif, who established the connection to the University of Auckland in New Zealand, Dr. Dominik Haneberg, who made the exchange possible and Andreas Angerer who helped me to find my way and write this thesis. I would also like to thank the Robotics Group at the University of Auckland under the supervision of Assoc. Prof. Bruce MacDonald, where I was able to work on the project. You inspired me with the work you do and were of great help whenever I needed it.

No words can express how thankful I am for the support of my parents Annemarie and Martin, and my sister Lisa. Thank you for your never ending support. I can count on you in good times and in bad.

There are so many people who helped me, it is hard to name them all. I thank you, if you ever had a coffee with me and talked with me about work and the life beyond work.

\newpage
\chapter*{Abstract}
Data collected during debugging is traditionally rendered as text. The special application field of robotics faces problems with this approach, since a typical robotic system constantly gathers and processes data from the surrounding environment through sensors. Developers of robotic applications are confronted with high amounts of data during debugging, which becomes hard to interpret if the data is represented as text and high amounts of data need to be interpreted at once. Robotic applications are also hard to interrupt during debugging, since robots generally do not run in a deterministic and suspendable environment.

This thesis tries to solve this problem by introducing a visual debugging system to support debugging of robotic applications: It takes into account the special requirements for a debugging tool in a robotic development environment, especially the uninterrupted rendering and the need for better visualization of debugging data to support a faster interpretation of this data. The goal of the developed system is to help developers understand the data during debugging more quickly and improve overall productivity during robot development. To achieve this goal, a system was designed and developed where developers can choose how they want to visualize data collected during debugging of a robotic application. This work presents the design and a prototypical implementation of the proposed system. Evaluating the hypothesis that the developed system can improve productivity during debugging is out of scope for this work but initial experiments were promising and the developed tool provides a solid base for further evaluations.
\newpage
\chapter*{TODOS}
\todo{listing positions}\\
\todo{dates on title pages}\\
\todo{general layout}\\
\newpage
\tableofcontents
\newpage
\listoffigures
\newpage
%\listoftables
\lstlistoflistings

\pagenumbering{arabic}
%end preface

\include{tex/introduction}
\include{tex/related}

\include{tex/visual_debugging_system}
\include{tex/rosdashboard}
\include{tex/casestudy}

\include{tex/conclusion}

\appendix
\include{tex/sources}
\end{document}
